---
title: "Activity Data"
output: md_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
This goal of this docment is to demonstrate a simple analysis of human activity data.

#Processing the Data

This analysis uses the Dplyr and the GGplot2 package that have to be loaded first
```{r message=FALSE, warning=FALSE}
library(dplyr)
library(ggplot2)
```


For reading the data it is assumed that the data is in the working directory.

```{r}
activityData<-read.csv("activity.csv")
```


A quick look at the data ensures that all the variables are in an approbriate form for analysis
```{r}
str(activityData)
```


#Exploratory analysis
For each day the sum of steps is formed. The Histogram shows the frequency of daily sums of steps

```{r}
stepData<- activityData  %>%group_by(date) %>% summarize(sum(steps))

hist(stepData$`sum(steps)`, breaks = 10)
```

The mean and median sum of steps across all days:

```{r}
median(stepData$`sum(steps)`, na.rm=TRUE )

mean(stepData$`sum(steps)`, na.rm=TRUE )
```
Mean and the median are very close together

#Activity patterns

Now we are examining the average number of steps in all intervals. A plot shows the 5 minute intervals and the average number of steps taken across days.

```{r}

stepDataInterval<- activityData  %>%group_by(interval) %>% summarize(AverageSteps=mean(steps, na.rm=TRUE))

ggplot(stepDataInterval, aes(x=interval, y=AverageSteps)) + geom_line()

```

The interval with the maximum average number of steps across all days is
calculated by the following code:

```{r}

maxIntervall<-stepDataInterval %>% filter(AverageSteps==max(AverageSteps, na.rm = TRUE))
maxIntervall[,1]

```
The finding is that interval 835 is the interval with the highest average number of steps.


#Dealing with missing values

To get an idea how many values are missing the number of rows are counted
that contain a missing value

```{r}
nrow(activityData  %>% filter(is.na(steps)))
```
The missing data shall be replaced by an assumption. In this case the average steps per interval across all days and intervals is taken for filling the gaps.
A new variable is assigned to create a new dataset without NA values.

```{r}
activityDataFilled<- activityData 
activityDataFilled<-replace(activityDataFilled,is.na(activityDataFilled), mean(activityData$steps, na.rm=TRUE))
stepDataFilled<- activityDataFilled  %>%group_by(date) %>% summarize(sum(steps))
```

We make the same Histogram for the number of steps as above in order to see what filling the missing data has changed

```{r}
hist(stepDataFilled$`sum(steps)`, breaks = 10)

median(stepDataFilled$`sum(steps)`, na.rm=TRUE )

mean(stepDataFilled$`sum(steps)`, na.rm=TRUE )
```


#Finding different activity patterns
Finally we analyse whether there are differences in activity patterns across weekends and week days? A new variable is introduced indicating whether the day is weekend or not.

```{r}
activityDataFilled$WE<-weekdays(as.Date(activityDataFilled$date))

activityDataFilled$WE<-ifelse(activityDataFilled$WE %in% c("Samstag","Sonntag"),"Weekend","Weekday")

stepDataAvFilled<- activityDataFilled  %>% group_by(interval, WE) %>% summarize(AverageSteps=mean(steps))

ggplot(stepDataAvFilled, aes(x=interval, y=AverageSteps)) + geom_line()+  
   facet_grid( WE ~.)


```
I can be seen from the Charts, that on weekends, the activity starts later in the day. An explanation would be that you sleep longer if you dont have to get up early for work.

